Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-13 paper/paper.tex
perfect partition and new sections

Formally, the online \emph{Balanced RePartitioning} problem (\OBRP{}) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
if they are assigned to the same cluster.
%\maciek{We can ommit the formal definition: "I.e.,
%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". %However, if we use the notion of cluster later, then TODO: define cluster}
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
Before serving the request,
the online algorithm
can re-partition,
i.e.,
it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
Once the re-partition is executed,
we serve
the communication request between two  nodes at cost~0 if they are collocated,
or at cost~1 if they are located in different clusters.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.

\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}

@@

\section{Perfect Partition Learner Algorithm}	\label{sec:upperbound}

@@

      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}

      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current partition $P$ with a (perfect) partition closest to $P_I$.
Hence it never moves to a partition beyond distance $\Delta$.      
\begin{property} \label{prop:dist<OPT}
    Let $P$ be any partition chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
    Then, $\mathit{dist}(P,P_I) \leq \Delta$.
\end{property}

@@

\begin{theorem}	\label{thm:upperbound}
    Algorithm \ref{alg:ppl} reaches the final partition $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
      On each inter-cluster request,
     the algorithm enumerates all $\ell$-way partitions of components
     that are in the same (closest) distance of $P_I$.
     That is, 
     once it reaches $\Delta = \mathit{dist} (P_{t}, P_I)$,
     it does not move to a partition
     $P_{t'}$, for any $t'>t,\mathit{dist} (P_{t'}, P_I) > \Delta$,
     before it enumerates all partitions at distance $\Delta$.
     Therefore,
     PPL eventually reaches $\Delta=\OPT$ and the partition $P_F$.
%    including the request that completes revealing of all components that are collocated in $P_F$.
    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{re\mhyphen partition}$
     (i.e., the number of internal edges in $P_F$).
    By Lemma \ref{lemma:rebalancecost},
    each re-partition costs at most $2\cdot\OPT$.
    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$ which implies the competitive ratio.
    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
    	So it is 4-competitive (?)}
 \end{proof}

\section{$O(\ell)$-competitive algorithm for $k=3$}

@@

>>>>

Formally, the online \emph{Balanced RePartitioning} problem (\OBRP{}) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
if they are assigned to the same cluster.
%\maciek{We can ommit the formal definition: "I.e.,
%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". %However, if we use the notion of cluster later, then TODO: define cluster}
%\maciek{We can ommit the formal definition: "I.e.,
%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
Before serving the request,
the online algorithm
can re-partition,
i.e.,
it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
Once the re-partition is executed,
we serve
the communication request between two  nodes at cost~0 if they are collocated,
or at cost~1 if they are located in different clusters.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.

\subsection{Perfect Partition Model}\label{sec:perfectPartition}
\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}

In this paper we mostly focus on a restricted variant of   \PPOBRP{}
where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
at the beginning and stays there permanently.
Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.

@@

\section{Recovering a Perfect Partition}	\label{sec:upperbound}

@@

%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}
      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current partition $P$ with a (perfect) partition closest to $P_I$.
Hence it never moves to a partition beyond distance $\Delta$.      
\begin{property} \label{prop:dist<OPT}
    Let $P$ be any partition chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
    Then, $\mathit{dist}(P,P_I) \leq \Delta$.
\end{property}

@@

\begin{theorem}	\label{thm:upperbound}
    Algorithm \ref{alg:ppl} reaches the final partition $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
      On each inter-cluster request,
     the algorithm enumerates all $\ell$-way partitions of components
     that are in the same (closest) distance of $P_I$.
     That is, 
     once it reaches $\Delta = \mathit{dist} (P_{t}, P_I)$,
     it does not move to a partition
     $P_{t'}$, for any $t'>t,\mathit{dist} (P_{t'}, P_I) > \Delta$,
     before it enumerates all partitions at distance $\Delta$.
     Therefore,
     PPL eventually reaches $\Delta=\OPT$ and the partition $P_F$.
%    including the request that completes revealing of all components that are collocated in $P_F$.
    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{re\mhyphen partition}$
     (i.e., the number of internal edges in $P_F$).
    By Lemma \ref{lemma:rebalancecost},
    each re-partition costs at most $2\cdot\OPT$.
    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$ which implies the competitive ratio.
%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
%    	So it is 4-competitive (?)}
 \end{proof}

\subsection{$O(\ell)$-competitive algorithm for $k=3$}

@@

\section{Arbitrary Communication Graph} \label{sec:generalModel}
